---
title: "Cleaning, Wrangling And Visualizing data on human exposure methods  ONTOX Systematic Scoping Review study"
author: "Maria Kalyva"
date: "`r Sys.Date()`"
output: word_document
header-includes:
   - \usepackage{booktabs}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage[table]{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, message = FALSE, warning = FALSE)
```

Import the .csv file downloaded from the Sysrev web platofm into R, with data on human exposure methods from the ONTOX Systematic Scoping Review study.

```{r import}
data <- read.csv("UserAnswers_A279.csv", header = TRUE, fill = TRUE, stringsAsFactors = FALSE)
class(data)
View(data)
```

Subset columns and keep those of interest, rename columns with the data labels of interest and discard one paper of wrong assessment.

```{r subset}
library(dplyr)
data <- subset(data, select = -c(Article.Title, Resolve., Include, 
                          Exposure_Model, Title, Journal,Authors))
data <- data %>% dplyr::rename("Programming_Statistical_analysis"  = "Programming...Statistical.analysis", 
  "scope_of_study"="scope.of.study", "Known_models"= "Known.models", "Exposure_Model_new" = "Exposure..Model...new", "Analysis_Method" = "Analysis..Method", "Population_study" = "Population.study")
names(data)  
data <- data[-c(669), ]
View(data)
```

In many cases Reviewers used the word *'Other'*, *'other'* or *'not reported'* when in the drop menu of options inside Sysrev, was not available the suitable information. They had the possibility to input the correct information in the Notes column. We need to remove these words as they are not actual information. The word *'equation'* was a wrong input of options for the Analysis method label at Sysrev, we have to remove it.

```{r remove}
data[data =="Other"] <- NA
data[data =="other"] <- NA
data[data =="not reported"] <- NA
data[data == "equations"] <- NA
View(data)
```

Unite data from the User.Note column with the scope.of.study column, as User.Note has input data regarding the scope of the study.

```{r unite}
library(tidyverse)
data <- data %>% 
  unite(scope_of_study, c(User.Note, scope_of_study), sep = ",")
data[data ==","] <- NA
View(data)
```

Fill the blank rows with NAs in all columns.

```{r fillNAs}
library(dplyr)
data <- data %>%
  mutate_at(c("Analysis_Method", "Chemical", "Exposure_Model_new", "Programming_Statistical_analysis",
              "Data", "Population_study", "Known_models", "scope_of_study",
              "Exposure_Route","Country", "Notes"), ~ na_if(., ''))
View(data)
```

For each paper the reviewing process was performed twice by two different Reviewers. In the first round the information was set/add on the labels of data extraction project at Sysrev. However, in the second reviewing as the Reviewers were assisted by a data extraction programming tool, they were just checking if the data extracted were correct, and only information that was missing was set/add, otherwise it was just left blank. The combined information of the two Reviewers is the complete data extracted for each paper, plus the (User.Name = tom+ontox-api),preformed on the stage Full-text screening extraction stage and contains important info on Country, Exposure_route and Exposure_Model. Each Reviewer inputs data in one row per article (Article.ID), therefore we need to paste the info of the 3 rows for each data label per article.

```{r paste}
library(dplyr)
data <- data %>% 
  group_by(Article.ID) %>% 
  summarise(across(everything(), ~ paste(., collapse = ";")))
View(data)
```

# Plot Year vs Analysis Methods

Separate the year_of_study variable into 3 columns that contain input data from the 3 Reviewers and coalesce them keep the fist non-missing value.

```{r cleanYear}
library(tidyr)
data <- data %>% 
  separate(year_of_study, c('yearI', 'yearII', 'yearIII'))
data[data == "NA"] <- NA
data <- data %>% 
  mutate_at(c("yearI", "yearII", "yearIII"), as.integer)
data$year_of_study <- coalesce(data$yearI, data$yearII, data$yearIII)
View(data)
```

Separate the 3 Reviewers input data of the Analysis_Method column. Some times the input data is added or just approved that is then inserted as a blank or NA. Coalesce the columns discarding NAs and keeping the strings that are not repeating by selecting the order.

```{r cleanAnalysis_Method}
library(tidyverse)
data <- separate(data=data, col = Analysis_Method, into = c("AnI", "AnII", "AnIII"), sep = ";")
data[data == "NA"] <- NA
data$AnI_AnII <- dplyr::coalesce(data$AnI, data$AnII)
data$Analysis_Method <- coalesce(data$AnIII, data$AnI_AnII)
View(data)
```

Some of the information is mentioned twice e.g. deterministic,statistical and statistical deterministic, we need to make this being one term as category so we can group them. The deterministic, statistical term means the exposure was done in a deterministic way and they include some statistics for correlation, so we correct this with deterministic. The model simulation is not an analysis method. It will not be representative here to keep it for the graph. We need the journals that had used as analysis method probabilistic, deterministic, and statistical analysis. I change the model simulations to deterministic, model simul., prob., to probabilistic etc. Usually, when using modelling tools, they aim for probabilistic studies.

```{r replaceString}
# Replace String with Another String on the Analysis_Method column
data <- data %>% 
  mutate(Analysis_Method = dplyr::recode(Analysis_Method, "equations,deterministic" = "deterministic", 
"deterministic,equations" = "deterministic","deterministic,probabilistic,statistical" = "deterministic,probabilistic",
"deterministic,probabilistic,equations" = "deterministic,probabilistic",
"probabilistic,deterministic" ="deterministic,probabilistic",
"probabilistic,deterministic,model simulations" = "deterministic,probabilistic",
"deterministic,statistical" = "deterministic", "probabilistic,statistical" = "probabilistic",
"probabilistic,statistical,model simulations" = "probabilistic","statistical,deterministic" = "deterministic",
        "statistical,probabilistic" = "probabilistic", "probabilistic,model simulations" = "probabilistic",
  "deterministic,model simulations,probabilistic" = "deterministic,probabilistic",
  "model simulations,statistical" = "statistical", 
"model simulations,deterministic,probabilistic" ="deterministic,probabilistic",
"model simulations,probabilistic" = "probabilistic","model simulations" = "deterministic,probabilistic"))
View(data)
```

Plot Exposure route vs Analysis methods

```{r plotYears}
yearly_counts_graph <- data %>%
  count(year_of_study, Analysis_Method) %>%
  ggplot(aes(x = as.numeric(year_of_study), y = n, color = Analysis_Method)) +
  geom_point(size = 3.5) +
 # geom_line() +
  geom_smooth() +
  scale_x_continuous(breaks = seq(2000, 2022, 5), 
                     limits = c(2000, 2022)) +
  scale_y_continuous(breaks = seq(1, 10, 3), 
                     limits = c(1, 10)) +
  theme_bw() +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 17),
        axis.text.y = element_text(face = 'bold', size = 17),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 22),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Yearly number of Analysis methods", #subtitle = "Plot counts vs years", 
       #caption = "ONTOX - Scoping review study",
       x = "Years", y = "Counts of Analysis methods", color = "Analysis methods")
yearly_counts_graph

```

# Plot Exposure route vs Analysis methods

```{r cleanEx.Route}
data <- separate(data=data, col = Exposure_Route, into = c("ExI", "ExII", "ExIII"), sep = ";")
library(stringr)
rep_str<-c("Dermal,Inhalation,Oral" = "Oral,Dermal,Inhalation", "Dermal,Oral" = "Oral,Dermal",
  "Dermal,Oral,Inhalation" = "Oral,Dermal,Inhalation","Inhalation,Dermal" = "Dermal,Inhalation",
  "Inhalation,Dermal,Oral" = "Oral,Dermal,Inhalation","Inhalation,Oral" = "Oral,Inhalation",
"Inhalation,Oral,Dermal" = "Oral,Dermal,Inhalation","Oral,Inhalation,Dermal" = "Oral,Dermal,Inhalation")
data$ExI<-str_replace_all(data$ExI, rep_str)
```

Plot Exposure route vs Analysis methods

```{r}
Exposure_analysis_plot <- data %>%
  ggplot(aes(x= Analysis_Method, fill = ExI)) +
  geom_bar(position = "fill") +
  scale_fill_discrete(name="Exposure route") +
  stat_count(geom = "text", 
           aes(label = stat(count)),
           position=position_fill(vjust=0.5), colour="white") +
theme_bw() +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 15),
        axis.text.y = element_text(face = 'bold', size = 15),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 22),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Exposure route versus Analysis methods", 
       #caption = "ONTOX - Scoping review study",
       x = "Analysis Method", y = "Exposure route counts in proportion")
Exposure_analysis_plot 
```

# Plot Exposure route vs Chemicals

```{r cleanChemColumn}
# We separate the data input by each one of the 3 Reviewers
data <- separate(data=data, col = Chemical, into = c("ChI", "ChII", "ChIII"), sep = ";")
data <- data %>%
  mutate_at(c("ChI", "ChII", "ChIII"), ~ na_if(., 'NA'))
data$ChI_ChII <- dplyr::coalesce(data$ChI, data$ChII)
data$Chemical<-coalesce(data$ChI_ChII, data$ChIII)
dataIII <-  data %>% 
  subset(select = c( Article.ID, Chemical, ExI, Exposure_Model_new)) 
#Fix the order of the Exposure route labels, so we are able to group them
library(stringr)
rep_str<-c( "Bisphenols,Non-persistent" = "Bisphenols","Cosmetic ingredients,Non-persistent" = "Cosmetic ingredients",
"VOCs,Other" = "VOCs","Cosmetic ingredients,Other" = "Cosmetic ingredients",
"Cosmetic ingredients,Persistent" = "Cosmetic ingredients","Metals,Other" = "Metals","Other,Metals" = "Metals",
"VOCs,Other" = "VOCs","Other,VOCs" = "VOCs","Non-persistent,Other" = "Non-persistent",
"Other,Non-persistent" = "Non-persistent","Pesticide,Non-persistent,Persistent" = "Pesticide",
"VOCs,Phthalates,Other" ="VOCs,Phthalates","Other,Pesticide" = "Pesticide",
"Metals,Mycotoxins,Persistent" = "Metals,Mycotoxins","Metals,Pesticide,Persistent" = "Metals,Pesticide",
"Metals,Persistent,Non-persistent,VOCs" = "Metals,VOCs","Persistent,PFASs" = "PFASs","PFASs,Persistent" = "PFASs",
"Pesticide,Metals" = "Metals,Pesticide","Persistent,Metals" = "Metals","Persistent,VOCs" = "VOCs",
"Pesticide,Persistent" = "Pesticide",
"Cosmetic ingredients,PFASs,Phthalates,Persistent" = "Cosmetic ingredients,Phthalates",
"Phthalates,Cosmetic ingredients" = "Cosmetic ingredients,Phthalates","Pesticide,VOCs,Cosmetic ingredients" = "Cosmetic ingredients,Pesticide,VOCs","VOCs" = "Volatile Org. Compounds",
"Pesticide" = "Pesticides", "Food additive" = "Food additives")
dataIII$Chemical <- str_replace_all(dataIII$Chemical, rep_str)
# We separate further as 1 Reviewer was possible to put more than one Chemical per paper
dataIII<-separate(data =dataIII, col = Chemical, into = c("CI", "CII","CIII","CIV"), sep = ",")
# We pivot longer to get 1 Chemical per article
dataIV <- dataIII %>% 
  pivot_longer(cols=c('CI', 'CII', "CIII", "CIV"),
  names_to='ChemicalC',
  values_to='Chemicals')%>%
  drop_na()
##Remove leading and trailing white space on each row
dataIV <-dataIV %>% 
  mutate_all(str_trim)
View(dataIV)

```

Plot Exposure route vs Chemicals with facets

```{r}
Ch_Exp_P <- dataIV %>% 
  group_by(Chemicals, ExI) %>% 
  summarize(n = n()) %>%
  mutate(perc = 100*n/sum(n)) %>%
  ggplot(aes(x = Chemicals, y = perc, fill = ExI)) +
  geom_bar(stat = "identity") +
  scale_fill_discrete(name="Exposure route") +
  #geom_col(show.legend = F) +
  facet_grid( ~ ExI, labeller = label_wrap_gen(width=20)) + 
  coord_flip() +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, 50))+
 # ylim(c(0, 116, 25)) +
  geom_text(aes(label = round(perc, 0)), size = 6, hjust = -.2) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(strip.text.x = element_text(face = 'bold', size = 25),
        strip.background = element_rect(fill="white", colour="black",size=1))  +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 15),
        axis.text.y = element_text(face = 'bold', size = 19),
        legend.text = element_text(size = 11),
        legend.title = element_text(size = 11),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 17),
        strip.text.x = element_text(size = 12), #change the facets font
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Chemicals versus Exposure route", 
       #caption = "ONTOX - Scoping review study",
        y = "percentage (%)", x = "Chemical classes")
Ch_Exp_P
```

# Estimation Methodologies versus Chemicals

```{r cleanEstMethodologies}
###separate the data on Estimation methodologies column from the 3 Reviewers
dataIV<-separate(data =dataIV, col = Exposure_Model_new, into = c("MI", "MII","MIII"), sep = ";")
dataIV <- dataIV %>%
  mutate_at(c("MI", "MII","MIII"), ~ na_if(., 'NA'))
#coalesce the columns
dataIV$Est_MI <- dplyr::coalesce(dataIV$MI, dataIV$MII)
dataIV$Est_M <- dplyr::coalesce(dataIV$Est_MI, dataIV$MIII)
# Replace String with Another String so we can group them. The label Script/Software,Mathematical model depends initially on a model, so Mathematical model is more appropriate and Script/Software,Toolbox is described better by a Toolbox.
library(dplyr)
dataIV <- dataIV %>% 
  mutate(Est_M = dplyr::recode(Est_M, "Script/Software,Mathematical model" = "Mathematical model",
"Script/Software,Empirical equations" = "Empirical equations","Script/Software,Toolbox" = "Toolbox/Software",
"Toolbox,Script/Software" = "Toolbox/Software", "Toolbox" = "Toolbox/Software",
"Mathematical model,Toolbox" = "Toolbox","Empirical equations,Script/Software" = "Empirical equations",
"Mathematical model,Script/Software" = "Mathematical model","Script/Software" = "Toolbox/Software"))
```

Plot Estimation Methodologies versus Chemicals

```{r}
library(forcats)
library(tidytext)
library(tidyverse)
library(tidyr)
Est_Chem_P <- dataIV %>% 
  group_by(Chemicals, Est_M) %>% 
  summarize(n = n()) %>%
  mutate(perc = round(100*n/sum(n))) %>%
  mutate(Est_M = as.factor(Est_M)) %>% 
  ggplot() +
  scale_fill_brewer(palette = "Set2") +
  geom_bar(aes(x = reorder_within(Chemicals, perc, Est_M),  
               y = perc, fill = Est_M), stat = "identity", show.legend = FALSE) +
  geom_text(aes(x = reorder_within(Chemicals, perc, Est_M), y = perc,
                label = round(perc, 1)), size =6, hjust = -.1,
                position = position_dodge(width = 1),
                          inherit.aes = TRUE) +
  facet_wrap(~ Est_M, scales = "free_y", labeller = label_wrap_gen(width=10)) +
  coord_flip() +
  scale_x_reordered() +
  scale_y_continuous(limits = c(0, 90)) +
 # ylim(c(0, 134)) +
  theme_bw() +
  theme(strip.text.x = element_text(face = 'bold', size = 17),
        strip.background = element_rect(fill="white", colour="black",size=1)) +
  theme(legend.position = "none") +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 15),
        axis.text.y = element_text(face = 'bold', size = 20),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 17),
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(size = 20))+
  labs(#title = "Estimatation Methodologies versus Chemicals", 
       # caption = "ONTOX - Scoping review study", 
       x = "Chemical classes", y = "percentage (%)")
Est_Chem_P

```

# Plot the number of studies for each study design

The study design shows what was the article's exposure calculation focus regarding population, for example, population, individual, age groups, cohort etc..

```{r cleanStudyDesign}
dataP <-  data %>% 
  subset(select = c( Article.ID, Population_study)) 
dataP<-separate(data =dataP, col = Population_study, into = c("PI", "PII","PIII"), sep = ";")
dataP[dataP == "NA"] <- NA
dataP$PI_PII <- dplyr::coalesce(dataP$PI, dataP$PII)
dataP$Pop_Anal <- coalesce(dataP$PI_PII, dataP$PIII)
library(stringr)
rep_str<-c( "cohort,population analysis" = "population analysis,cohort",
"individual,population analysis" = "population analysis,individual","age groups,cohort" = "cohort,age groups",
            "population analysis,age groups,individual" = "population analysis,age groups",
            "age groups,population analysis" = "population analysis,age groups",
"individual,age groups" = "age groups,individual","individual,cohort" = "cohort,individual")
dataP$Pop_Anal <- str_replace_all(dataP$Pop_Anal, rep_str)
```

Plot number of studies per study design

```{r}
Pop_anal_plot <- dataP %>%
  group_by(Pop_Anal) %>% 
  summarize(n = n()) %>%
  mutate(perc = (round(100*n/sum(n), 1))) %>%
 # count() %>% 
  ggplot(aes(x = reorder(Pop_Anal, +perc), y = perc)) +
  geom_bar(stat = "identity", color = "darkblue", fill = 'lightblue') +
  coord_flip() +
  geom_text(aes(label = perc), 
            hjust = -.2) +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 12),
        axis.text.y = element_text(face = 'bold', size = 20),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        strip.text.x = element_text(size = 12), #change the facets font
        plot.caption = element_text(face = 'bold', size = 10),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Types of study design", 
       #caption = "ONTOX - Scoping review study",
       x = "Types of study design", y = "percentage (%)")
Pop_anal_plot
```

# Plot number of studies per scope of each paper

The scope of the study column contains information about what is the focus of the paper, e.g. to calculate human exposure, risk, human exposure and liver, kidney, brain or other health outcomes.

```{r}
#Subset the data of interest
dataS <-  data %>% 
  subset(select = c( Article.ID, scope_of_study)) 
#Separate the data per Reviewer
dataS<-separate(data =dataS, col = scope_of_study, into = c("SI", "SII","SIII"), sep = ";")
dataS[dataS == 'NA'] <- NA
#coalesce
dataS$SI_SII <- dplyr::coalesce(dataS$SI, dataS$SII)
#Some Reviewers had more than one input, in order to get one information per article we unite and then separate the data
dataS <- dataS %>%  
  unite(col = "Sc", c("SI_SII", "SIII"), sep = ",") %>% 
  separate(col = "Sc", into = c("S1","S2","S3","S4", "S5"), sep = ",")
dataS[dataS == 'NA'] <- NA
dataS[dataS == ''] <- NA
dataS1 <- dataS %>% 
  pivot_longer(cols=c("S1","S2","S3","S4"),
               names_to="Scs",
               values_to="Scope")
#Remove duplicates on selected columns, recode empty stings by NAs and remove them 
dataS1 <- dataS1 %>% 
  distinct(Article.ID, Scope, .keep_all = TRUE)
View(dataS1)

###subset the column of interest
dataS1 <-  dataS1 %>% 
  subset(select = c( Article.ID, Scope)) %>% 
  na.omit()
View(dataS1)

#We remove leading and trailing white space on each row
dataS1 <-dataS1 %>% 
  mutate_all(str_trim)
View(dataS1)

##### Replace String with Another String on the Known models column
library(dplyr)
dataS1 <- dataS1 %>% 
  mutate(Scope = dplyr::recode(Scope, "human exposure (HE) estimates only" = "Human exposure estimates",
"human exposure (HE) estimates onlyNA" = "Human exposure estimates",
"HE associated with risk assessment"= "Human exposure and risk assessment",
"HE associated with risk assessmentNA"= "Human exposure and risk assessment",
"HE associated with other health outcomes" = "Human exposure and health outcomes",
"HE associated with liver toxicity" = "Human exposure and liver toxicity",
"HE associated with brain toxicity" = "Human exposure and brain toxicity",
              "HE associated with kidney toxicity" = "Human exposure and kidney toxicity",
"HE associated with kidneys toxicity" = "Human exposure and kidney toxicity",
"validation biomonitoring" = "Human exposure and biomonitoring validation",
          "Validation againist  DEEM-FCID Version 3.16 software used by EPA USA" = "Human exposure and biomonitoring validation"))
View(dataS1)
```

Plot number of studies per scope of the paper

```{r}
Scope_plot <- dataS1 %>% 
  group_by(Scope) %>% 
  summarize(n = n()) %>%
  mutate(perc = (round(100*n/sum(n), 0))) %>% 
  ggplot(aes(x = reorder(Scope, +perc), y = perc)) + 
  coord_flip()+
  geom_text(aes(label = perc), 
            hjust = -.2) +
  geom_bar(stat = "identity", color ="blue", fill = 'orange')+
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 14),
        axis.text.y = element_text(face = 'bold', size = 20),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Scope of the papers", 
       #caption = "ONTOX - Scoping review study",
       x="Scope of study", y = "percentage (%)")
Scope_plot

```

# Cleaning and reshaping the data of the Notes column

The Notes column contains a combination of data from the other data label columns. When the information was not offered by the drop menu at the Sysrev web platform, the Reviewers had to write down on the Notes column, the name of the data label column a semicolon (:) and the the information

```{r cleanNotesColumn}
###subset the column of interest
dataN <-  data %>% 
  subset(select = c( Article.ID, Notes)) 
View(dataN)

###separate each row of the Notes column at the symbol ';' , put in from the 3 reviewers
dataN<-separate(data =dataN, col = Notes, into = c("NI", "NII","NIII", "NIV", "NV", "NVI"), sep = ";")
View(dataN)

###Substitute "NA" with NA
dataN[dataN == 'NA'] <- NA
View(dataN)

# Remove leading and trailing white space on each row
dataN <-dataN %>% 
  mutate_all(str_trim)
View(dataN)

# Separate the all the columns to Names and Values
library(tidyr)
dataN <-separate(data =dataN, col = NI, into = c("NamesI", "ValuesI"), sep = ":")
dataN <-separate(data =dataN, col = NII, into = c("NamesII", "ValuesII"), sep = ":")
dataN <-separate(data =dataN, col = NIII, into = c("NamesIII", "ValuesIII"), sep = ":")
dataN <-separate(data =dataN, col = NIV, into = c("NamesIV", "ValuesIV"), sep = ":")
dataN <-separate(data =dataN, col = NV, into = c("NamesV", "ValuesV"), sep = ":")
dataN <-separate(data =dataN, col = NVI, into = c("NamesVI", "ValuesVI"), sep = ":")
View(dataN)

####pivot_wider the NI,NII, NIII, NIV, NV, NVI columns
library(tidyr)
dataNI<- dataN %>% 
  subset(select = c( Article.ID, NamesI, ValuesI)) %>% 
  pivot_wider(names_from = NamesI, values_from = ValuesI) %>% 
  select(-'NA')
View(dataNI)

dataNII<- dataN %>% 
  subset(select = c( Article.ID, NamesII, ValuesII)) %>% 
  pivot_wider(names_from = NamesII, values_from = ValuesII)%>% 
  select(-'NA')
View(dataNII)

dataNIII<- dataN %>% 
  subset(select = c( Article.ID, NamesIII, ValuesIII)) %>% 
  pivot_wider(names_from = NamesIII, values_from = ValuesIII)%>% 
  select(-'NA')
View(dataNIII)

dataNIV<- dataN %>% 
  subset(select = c( Article.ID, NamesIV, ValuesIV)) %>% 
  pivot_wider(names_from = NamesIV, values_from = ValuesIV)%>% 
  select(-'NA')
View(dataNIV)

dataNV<- dataN %>% 
  subset(select = c( Article.ID, NamesV, ValuesV)) %>% 
  pivot_wider(names_from = NamesV, values_from = ValuesV)%>% 
  select(-'NA')
View(dataNV)

dataNVI<- dataN %>% 
  subset(select = c( Article.ID, NamesVI, ValuesVI)) %>% 
  pivot_wider(names_from = NamesVI, values_from = ValuesVI)%>% 
  select(-'NA')
View(dataNVI)
```

# Create a new dataframe by joining the main dataframe with the dataframes from the Notes column

```{r}
# subset the Known_models and Data column from the main data dataframe
data_Mod_Dat <- data %>% 
  subset(select = c(Article.ID, Known_models, Data))
View(data_Mod_Dat)

# convert the data_Mod_Dat$Article.ID column as character, 
data_Mod_Dat$Article.ID <- as.character(data_Mod_Dat$Article.ID)    
class(data_Mod_Dat$Article.ID)

##### Join multiple data.frames
list_df = list(data_Mod_Dat,dataNI,dataNII,dataNIII,dataNIV,dataNV,dataNVI)
df_Tot <- list_df %>% reduce(full_join, by='Article.ID')
View(df_Tot)
```

Clean

```{r}
##Unite, separate and clean from NA and duplicated values total info column Known_modelsT 
library(tidyr)
df_Tot <- df_Tot %>%  
unite(col = "Known_modelsT", c("Known_models", "Known models.x", "Known models.y", "Known models.x.x", "Known models.y.y"),
      sep = ";") %>% 
  separate(col = "Known_modelsT", into = c("Km1","Km2","Km3","Km4","Km5","Km6","Km7"), sep = ";")
View(df_Tot)

###Substitute "NA" and "other" word with NA 
df_Tot[df_Tot == 'NA'] <- NA
View(df_Tot)

##### Unite all the information while removing NAs
library(tidyr)
df_Totfm <- df_Tot %>%  
  unite(col = "Known_modT", c("Km1","Km2","Km3","Km4","Km5","Km6","Km7"), sep = ",", na.rm = TRUE) %>% 
  separate(col = "Known_modT", into = c("K1","K2","K3","K4","K5","K6","K7"), sep = ",") 
View(df_Totfm)

df_Totfm[df_Totfm == "other"] <- NA

df_Totfm1 <- df_Totfm %>% 
  subset(select = c("Article.ID","K1","K2","K3","K4","K5","K6","K7"))
View(df_Totfm1)

df_Totfm2 <- df_Totfm1 %>% 
  pivot_longer(cols=c("K1","K2","K3","K4","K5","K6","K7"),
               names_to="Ks",
               values_to="Knownmod") %>%
                drop_na()
View(df_Totfm2)

#Remove duplicates on selected columns, fill empty stings by NAs and remove them
df_Totfm3 <- df_Totfm2 %>% 
  distinct(Article.ID,Knownmod, .keep_all = TRUE) %>% 
  mutate_at("Knownmod", ~ na_if(.,'')) %>% 
  na.omit()%>% 
  mutate_all(str_trim)
View(df_Totfm3)

###Remove the row 75 as it contains repeated information 
df_Totfm3 <- df_Totfm3[-c(75), ]
View(df_Totfm3)

##### Replace String with Another String on the Known models column
library(dplyr)
df_Totfm3 <- df_Totfm3 %>% 
  mutate( Knownmod = dplyr::recode( Knownmod, 'PRIMo (EFSA)' = 'Pesticide Residue Intake Model (PRIMo) EFSA', 
                        'PRIMo' = 'Pesticide Residue Intake Model (PRIMo) EFSA',
              'PRIMo (Pesticide Residue Intake Model)' = 'Pesticide Residue Intake Model (PRIMo) EFSA',
              'Cares' = 'Cumulative and Aggregate Risk Evaluation System (Cares) USEPA', 'CARES' = 'Cumulative and Aggregate Risk Evaluation System (Cares) USEPA', 
              'CalTox' = 'Multimedia transport & transform. model (CalTOX) USEPA', 'Caltox' = 'Multimedia transport & transform. model (CalTOX) USEPA', 
'CalTOX' = 'Multimedia transport & transform. model (CalTOX) USEPA',
                              'PRoTEGE' = 'PRoTEGE (Prioritiz./Ranking Toxic Exposures, GIS Ext.) eohsi US',
'Prioritization/Ranking of Toxic Exposures with GIS Extension (PRoTEGE)' = 'PRoTEGE (Prioritiz./Ranking Toxic Exposures, GIS Ext.) eohsi US',
          'Kinetic Dietary Exposure Model' = 'Kinetic Dietary Exposure Model (KDEM) INRA – Met@risk FR',
  'Kinetic Dietary Exposure Model (KDEM)' = 'Kinetic Dietary Exposure Model (KDEM) INRA – Met@risk FR',
  'Creme Food Safety' = 'Creme Food Safety® software Creme global',
    'Creme Food Safety® Software' = 'Creme Food Safety® software Creme global',
          'Creme RIFM model' = 'Creme RIFM Exposure Model Creme global',
      'Creme RIFM Exposure Model' = 'Creme RIFM Exposure Model Creme global',
              'FACET  (Flavours' = 'Flavours Additives and Food Contact Materials Exp. Task FACET FSA UK)',
                                    'SHEDS-HT' = 'High throughput Stoch. Human Exp. Dose Simul. model (SHEDS-HT) USEPA',
  'High throughput (HT) Stochastic Human Exposure and Dose Simulation model (SHEDS-HT)' = 'High throughput Stoch. Human Exp. Dose Simul. model (SHEDS-HT) USEPA',
'SHEDS-PM' = 'Population Exposure model (SHEDS-PM)', 'ExpoFIRST' = 'Exposure Factors Interactive Resource Scenarios Tool (ExpoFIRST) USEPA',
'Exposure Factors Interactive Resource for Scenarios Tool (ExpoFIRST)' = 'Exposure Factors Interactive Resource Scenarios Tool (ExpoFIRST) USEPA',
                              'PACEM' = 'Probabilistic Aggreg. Consm. Exp. Model (PACEM) RIVM',
                              'Dermal (PACEM-KD)' = 'Dermal Kinetic Model (PACEM-KD) RIVM',
                            'Probabilistic Aggregated Consm. Exposure Model–Kinetic' = 'Dermal Kinetic Model (PACEM-KD) RIVM',
                            'ECETOC TRA' = 'Targeted Risk Assessment tool (Ecetoc TRA)',
                              'Ecetoc Targeted Risk Assessment (Ecetoc TRA)' = 'Targeted Risk Assessment tool (Ecetoc TRA)',
                              'Sprayexpo' = 'SprayExpo Fraunhofer ITEM',
                              'SprayExpo' = 'SprayExpo Fraunhofer ITEM',
'Exposure Forecasting (ExpoCast)' = 'Exposure Forecasting (ExpoCast) USEPA', 'ExpoCast EPA' = 'Exposure Forecasting (ExpoCast) USEPA',
            'European Solvents Industry Group (ESIG) Generic Exposure Scenario (GES) Risk and Exposure Tool (EGRET)' = ' Risk and Exposure Tool (EGRET) ESIG/EU',
                          'EGRET' = 'Risk and Exposure Tool (EGRET) ESIG/EU',
       'USEtox' = 'Environment Program (USEtox) model UN','Hazardous Air Pollutant Exposure Model (HAPEM)' = 'Hazardous Air Pollutant Exposure Model (HAPEM) USEPA',
       'HAPEM or Hazardous Air Pollutant Exposure Model' = 'Hazardous Air Pollutant Exposure Model (HAPEM) USEPA',
       'The RAIDAR-Indoor and Consumer Exposure' = 'Risk Assess. Ident. And Ranking model (RAIDAR) cemc-TrentU CA',
          'Multimedia Exposure Assessment Model by USEPA' = 'Multimedia Exp. Assessment Model (MULTIMED) USEPA', 
       'EFAST-CEM' = 'Exposure & Fate Assess. Screening Tool (EFAST) USEPA',
        'Exposure and Fate Assessment Screening Tool (EFASTv2)' = 'Exposure & Fate Assess. Screening Tool (EFAST) USEPA',
         'RACE' = 'Rapid Assessment of Contaminant Exposure (RACE) EFSA',
                                    'FEIM' = 'Food Enzyme Intake Model (FEIM) EFSA', 
                                    'FAIM' = 'Food Additives Intake Model (FAIM) EFSA',
                                    'FACE' = 'Feed Additive Consumer Exposure (FACE) EFSA', 
                                    'REACT' = 'Reach Exposure Assessment Consumer Tool (REACT) AISE',
                                    'ABICAM' = 'Activity-Based Indoor Chemical Assessment Model (ABICAM) USEPA',
                                    'Pangea' = 'Multimedia fate and exposure assessment model (Pangea) SPH US',
                                    "Qsar(for risk assessment)" = "Open Qsar App (OPERA) NICEATM/USEPA",
                                    "OPERA" = "Open Qsar App (OPERA) NICEATM/USEPA",
                                    "ExDoM" = "Exposure Dose Model (ExDoM) Dept. Environmental Engg. UoC",
                                    "DERMWIN" = "Dermal permeability modeling program (DERMWIN) USEPA",
       "Combined Environmental Stressors' Exposure (CENSE) tool" = "Combined Environmental Stressors' Exp. (CENSE) tool AUTH",
                                    "Browse model" = "Bystanders Residents Operators WorkerS Exp. model BROWSE ssau UK",
                                    "Mentor software" = "Mentor tool eohsi US",
                                    "GExFRAME/Scibin" = "GExFRAME/Scibin EC US",
                                    'ArcGIS software' = 'ArcGIS software Esri',
      'BREAM (Bystander and Resident Exposure Assessment Model)' = 'Bystander and Resident Exposure Assess. Model (BREAM) fera UK',
                                    'VLIER-HUMAAN' = 'VLIER-HUMAAN Vito BE',
                                    'Far-field Human Exposure (FHX)' = 'Far-field Human Exposure (FHX) cemc-TrentU CA',
                                    'ICECRM' = 'Indoor chemical exp. classif./ranking model (ICECRM) UToronto CA',
                                    'IMPACT 2002' = "IMPACT 2002 epfl CH",
                                    'Modul’ERS tool' = 'Modul’ERS tool ineris FR',
                                    'INDAIR/EXPAIR' ='INDAIR/EXPAIR sei UK',
               'Transient Transport through the epiDERMis (TTDERM)' = 'Transient Transport through epiDERMis (TTDERM) lbl US',
                                    'PROduction-To-EXposure  (PROTEX)' = 'PROduction-To-EXposure  (PROTEX) UToronto CA',
                                    'positive matrix factorization (PMF)' = 'positive matrix factorization (PMF) USEPA',
                                    'personal delivered dose (PDD) model' = 'personal delivered dose (PDD) model BU SPH US',
                                    'INTEGRA web platform' = 'INTEGRA web platform ugl AU',
'Consumer Product Exposure and Risk assessment system (COPER)' = 'Consumer Product Exp. and Risk assess. (COPER) Yoon Seong-kyu KR',
                                    'Exposure Potential Model (DEPM)' = 'Exposure Potential Model (DEPM) USEPA',
  'ACC-HUMAN' = 'ACC-HUMAN itm su SE',
                                'Calendex EPA' = 'Calendar-based multi-pathway exp. model (DEEM-FCID/Calendex) USEPA',
                                  'IEUBK EPA' = "Integrated Exposure Uptake Biokinetic IEUBK USEPA",
                                  'ConsExpo RIVM' = 'Consumer Exposure & Uptake Model (ConsExpo) RIVM',
                                  'MCRA RIVM/EFSA' = 'Monte Carlo Risk Assessment (MCRA) RIVM/EFSA',
                                  'MERLIN-Expo LIFE-VERMEER project Italy' ='Modelling Exp. to chemic. for Risk assess. (MERLIN-Expo) EU',
                              'SHEDS-multimedia EPA' = 'Stoch. Human Exp. Dose Simul. model (SHEDS-multimedia) USEPA'))
View(df_Totfm3) 

####The Creme food global is a commercial software, that is we have to buy
###Remove the row 75 as it contains repeated information 
df_Totfm3 <- df_Totfm3[df_Totfm3$Knownmod != 'Creme Food Safety® software Creme global', ] 
df_Totfm3 <- df_Totfm3[df_Totfm3$Knownmod != 'Creme RIFM Exposure Model Creme global', ] 
df_Totfm3 <- df_Totfm3 %>% 
  mutate_all(str_trim)
View(df_Totfm3)



# dataIV contains clean the Exposure route, Chemicals and Estimated methodologies data columns 
View(dataIV)
###we tranform dataIV$Article.ID into character 
dataIV$Article.ID <- as.character(dataIV$Article.ID)

#We subset the Exposure route, Chemicals and Estimated methodologies columns that have been cleaned
dataIV_m <- dataIV %>% 
  subset(select = c(Article.ID, ExI, Chemicals, Est_M))
View(dataIV_m)

#Join the df df_Totfm3and dataIV_m that contains clean Known models, Chemicals, Exposure_route and Est_m data
df_Models <- full_join(df_Totfm3, dataIV_m, by = "Article.ID", multiple = "all", relationship = "many-to-many")
View(df_Models)


 
 
```

# Plot Known models versus Exposure route

```{r}
###Plot all the Known models found in our study versus Exposure route
Models_ExpRoute<- df_Models %>% 
 distinct(Article.ID,Knownmod,ExI, .keep_all = TRUE) %>% 
  na.omit() %>%
  group_by(Knownmod) %>% 
  count(ExI, Knownmod) %>% 
  ggplot(aes(y= Knownmod, x=n, fill= ExI)) +
  geom_bar(stat = "identity") +
  scale_fill_discrete(name="Exposure route") +
  theme_bw() +
  theme(axis.title = element_text(size = 15),
        axis.text.x = element_text(face = 'bold', size = 11),
        axis.text.y = element_text(face = 'bold', size = 9),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        #strip.text.x = element_text(size = 12), #change the facets font
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(title = "Known models versus Exposure route", 
       caption = "ONTOX - Scoping review study",
       x = "Counts of Exposure route", y = "Known models")
Models_ExpRoute
```

The models are too many, we filter the most frequent used models. Calculate the percentage within a group.

```{r}
library(formattable)
KnownModels_Per<- df_Models %>% 
  group_by(Knownmod) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq_per = formattable::percent(cnt / sum(cnt))) %>% #### relative frequency expressed in percentages
  arrange(desc(freq_per))
View(as.data.frame(KnownModels_Per))
```

Filter the most frequent models

```{r}
freq_knownMod <- c ('Stoch. Human Exp. Dose Simul. model (SHEDS-multimedia) USEPA', 'Monte Carlo Risk Assessment (MCRA) RIVM/EFSA',
      'Environment Program (USEtox) model UN', 'Consumer Exposure & Uptake Model (ConsExpo) RIVM',
    'Multimedia transport & transform. model (CalTOX) USEPA', 
    'Pesticide Residue Intake Model (PRIMo) EFSA', 'Exposure Forecasting (ExpoCast) USEPA',
  'Modelling Exp. to chemic. for Risk assess. (MERLIN-Expo) EU', 'Probabilistic Aggreg. Consm. Exp. Model (PACEM) RIVM',
  "Risk and Exposure Tool (EGRET) ESIG/EU", "Risk Assess. Ident. And Ranking model (RAIDAR) cemc-TrentU CA", 
  'Targeted Risk Assessment tool (Ecetoc TRA)')
df_Models_filt <- df_Models %>% 
  filter(Knownmod %in% freq_knownMod)
View(df_Models_filt)


#Recode some longer labels
df_Models_filt <- df_Models_filt %>% 
  mutate( Knownmod = dplyr::recode( Knownmod, 'Stoch. Human Exp. Dose Simul. model (SHEDS-multimedia) USEPA' = 'Stoch. Human Exp. Dose Simul. \nmodel (SHEDS-multimedia) USEPA',
                      'Monte Carlo Risk Assessment (MCRA) RIVM/EFSA' = 'Monte Carlo Risk Assessment\n(MCRA) RIVM/EFSA',                 
                      'Environment Program (USEtox) model UN' = 'Environment Program\n(USEtox) model UN',
                      'Consumer Exposure & Uptake Model (ConsExpo) RIVM' = 'Consumer Exposure & Uptake\nModel (ConsExpo) RIVM',
                      'Multimedia transport & transform. model (CalTOX) USEPA' = 'Multimedia transport & transform.\nmodel (CalTOX) USEPA',
                  'Pesticide Residue Intake Model (PRIMo) EFSA' ='Pesticide Residue Intake\nModel (PRIMo) EFSA',
                  'Exposure Forecasting (ExpoCast) USEPA' = 'Exposure Forecasting\n(ExpoCast) USEPA',
              'Modelling Exp. to chemic. for Risk assess. (MERLIN-Expo) EU' = 'Modelling Exp. to chemic. for\nRisk assess. (MERLIN-Expo) EU',
                  'Probabilistic Aggreg. Consm. Exp. Model (PACEM) RIVM' = 'Probabilistic Aggreg. Consm. Exp.\nModel (PACEM) RIVM',
                  "Risk and Exposure Tool (EGRET) ESIG/EU" = "Risk and Exposure Tool\n(EGRET) ESIG/EU",
        "Risk Assess. Ident. And Ranking model (RAIDAR) cemc-TrentU CA" = "Risk Assess. Ident. And Ranking \nmodel (RAIDAR) cemc-TrentU CA",
              'Targeted Risk Assessment tool (Ecetoc TRA)' = 'Targeted Risk Assessment\ntool (Ecetoc TRA)'))
View(df_Models_filt)

```

Plot Known models versus Exposure route

```{r}
Models_ExpRoute_filt<- df_Models_filt %>% 
   group_by(Knownmod, ExI) %>% 
  summarize(n = n()) %>%
  mutate(perc = round(100*n/sum(n), 0)) %>%
  ggplot(aes(y = Knownmod, x = perc, fill= ExI)) +
  geom_bar(stat = "identity") + 
 # scale_x_continuous(breaks = seq(0, 100, 25), 
         #            limits = c(0, 100)) +
  geom_text(aes(label = perc), size=6,
            position=position_stack(vjust=0.5), colour="white") +
  scale_fill_discrete(name="Exposure route") +
  theme_bw() +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 15),
        axis.text.y = element_text(face = 'bold',size = 18),
        legend.text = element_text( size = 18),
        legend.title = element_text(size = 20),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(face = 'bold',size = 20),
        plot.caption = element_text(face = 'bold', size = 15),
        plot.subtitle = element_text(face = 'bold',size = 15))+
  labs(#title = "Known models versus Exposure route", 
       #caption = "ONTOX - Scoping review study",
       x = "Percentage (%)", y = "Mathematical models and toolboxes")
Models_ExpRoute_filt
```

# Plot all the Known models found versus Chemicals

```{r}
Models_Chemicals<- df_Models %>% 
   distinct(Article.ID,Knownmod,Chemicals, .keep_all = TRUE) %>% 
  na.omit() %>%
  group_by(Knownmod) %>% 
  count(Chemicals, Knownmod) %>% 
  ggplot(aes(y= Knownmod, x=n, fill= Chemicals)) +
  geom_bar(stat = "identity") +
  scale_fill_discrete(name="Chemical classes") +
  theme_bw() +
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 12),
        axis.text.y = element_text(face = 'bold', size = 14),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        plot.caption = element_text(face = 'bold', size = 10),
        plot.subtitle = element_text(size = 20))+
  labs(#title = "Known models versus Chemical classes", 
       #caption = "ONTOX - Scoping review study",
       x = "Counts of chemical classes", y = "Mathematical models/toolboxes")
Models_Chemicals
```

## Plot only the 12 most used Known models versus Chemicals

```{r}
##Recode to short acronyms of the models
library(dplyr)
df_Models_filt_short <- df_Models_filt%>%
  mutate(Knownmod = dplyr::recode(Knownmod, "Monte Carlo Risk Assessment\n(MCRA) RIVM/EFSA" = "MCRA",
                                  "Consumer Exposure & Uptake\nModel (ConsExpo) RIVM" = "ConsExpo",
                                  "Stoch. Human Exp. Dose Simul. model\n (SHEDS-multimedia) USEPA" = "SHEDS\nmultimedia",
                                  'Multimedia transport & transform.\nmodel (CalTOX) USEPA' = 'CalTOX',
                                  'Pesticide Residue Intake\nModel (PRIMo) EFSA' = 'PRIMo',
                                  'Targeted Risk Assessment\ntool (Ecetoc TRA)' = 'Ecetoc',
                                  'Risk and Exposure Tool\n(EGRET) ESIG/EU' = 'EGRET',
                                  'Environment Program\n(USEtox) model UN' = 'USEtox',
                                  'Risk Assess. Ident. And Ranking model \n(RAIDAR) cemc-TrentU CA' = 'RAIDAR',
                                  'Probabilistic Aggreg. Consm. Exp.\nModel (PACEM) RIVM' = 'PACEM',
                                  'Exposure Forecasting\n(ExpoCast) USEPA' = 'ExpoCast',
                                  'Modelling Exp. to chemic. for\nRisk assess. (MERLIN-Expo) EU' = 'MERLIN-Expo'))
View(df_Models_filt_short)


##### Replace String with Another String on the Known models column
library(dplyr)
df_Models_filt_short <- df_Models_filt_short %>% 
  mutate(Chemicals = dplyr::recode(Chemicals, 'Pharmaceuticals' = 'Pharms.',
                                   "Non-persistent" = "Non persistent",
                                   "Enviromental emissions" = "Enviromen. emissions",
                          "Volatile Org. Compounds" = "Volatile Org. Comps.",
                          "Cosmetic ingredients" = "Cosmetic ingrs."))

#Plotting with faceting
Models_Chemicals_f <- df_Models_filt_short %>% 
  distinct(Article.ID,Knownmod,Chemicals, .keep_all = TRUE) %>% 
  na.omit() %>% 
  group_by(Knownmod, Chemicals) %>% 
  summarize(n = n()) %>%
  mutate(perc = 100*n/sum(n)) %>%
  ggplot(aes(x = Knownmod, y = perc, fill = Chemicals)) +
  geom_bar(stat = "identity") +
 # scale_fill_discrete(name="Chemicals") +
  geom_col(show.legend = F) +
  facet_grid( ~ Chemicals, labeller = label_wrap_gen(width=10)) + 
  coord_flip() +
  scale_y_continuous(limits = c(0, 160), breaks = seq(0, 160, 60))+
  geom_text(aes(label = round(perc, 0)), size = 6, hjust = -.1) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(strip.text.x = element_text(face = 'bold', size = 12),
        strip.background = element_rect(fill="white", colour="black",size=1)) + #change the facets font)
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 14),
        axis.text.y = element_text(face = 'bold', size = 17),
        #legend.text = element_text(size = 11),
        #legend.title = element_text(size = 11),
        #legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        plot.caption = element_text(face = 'bold', size = 10),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Known models versus Chemicals", 
       #caption = "ONTOX - Scoping review study",
       x = "Mathematical models / toolboxes", y = "percentage (%)")
Models_Chemicals_f
```

# Plot parameters (data input)  versus filtered Known models

Unite all the columns with the input data Data columns into a total info column DataT coming either from the Sysrev menu or as an input at the Notes colum by the Reviwers and separate to clean from NAs and duplicated values

```{r}
library(tidyr)
df_Tot_D <- df_Tot %>%  
  unite(col = "DataT", c("Data.x", "Data.y","Data.x.x", "Data.y.y","Data.x.x.x", "Data.y.y.y","Data"),
        sep = ";", na.rm = TRUE) %>% 
  separate(col = "DataT", into = c("D1","D2","D3","D4","D5","D6","D7"), sep = ";")

df_Tot_D[df_Tot_D == 'NA'] <- NA

##### Unite all the information with "," while removing NAs and separate by "," all the possible word inputs
library(tidyr)
df_Tot_Dm <- df_Tot_D %>%  
  unite(col = "Data", c("D1","D2","D3","D4","D5","D6","D7"), sep = ",", na.rm = TRUE) %>% 
  separate(col = "Data", into = c("d1","d2","d3","d4","d5","d6","d7"), sep = ",") 

##Remove the word "Other", it was used when the Reviewers had to input words at the Notes column
df_Tot_Dm[df_Tot_Dm == "Other"] <- NA

df_Tot_Dm1 <- df_Tot_Dm %>% 
  subset(select = c("Article.ID","d1","d2","d3","d4","d5","d6","d7"))
View(df_Tot_Dm1)

df_Tot_Dm2 <- df_Tot_Dm1 %>% 
  pivot_longer(cols=c("d1","d2","d3","d4","d5","d6","d7"),
               names_to="Ks",
               values_to="Data") %>%
  drop_na()

#Remove duplicates on selected columns, recode empty stings by NAs and remove them 
df_Tot_Dm3 <- df_Tot_Dm2 %>% 
  distinct(Article.ID, Data, .keep_all = TRUE) %>% 
  mutate_at("Data", ~ na_if(.,'')) %>% 
  na.omit()


##### Replace String with Another String on the Data column
library(dplyr)
df_Tot_Dm3 <- df_Tot_Dm3 %>% 
  mutate( Data = dplyr::recode( Data, 'area' = 'surface area', 'cross-sectional area' = 'surface area', 
'dermal absorption' = 'dermal absorption coefficient', 'fraction absorbed in skin' = 'absorbed fraction',
'absorption fraction' = 'absorbed fraction', 'dermal absorption efficiency' = 'absorption efficiency',
'area' = 'surface area', "hand area" = "surface area",
"skin surface area" = "surface area", "skin area" = "surface area",
  "average time" = "average exposure time", "averaging time" = 'average exposure time',
'avearing time' = 'average exposure time','averaged time' = 'average exposure time', 'time' = 'average exposure time',
    'time activity' = 'activity paterns(time)', 'partition coefficients' = 'partition coefficient',
  'skin permeability coefficient' = 'permeability coefficient', 'quantity of airborne spray' = 'quantity of spray',
'quantity of spray liquid' = 'quantity of spray', 'vapor pressure' = 'vapour pressure',
 'frequency' = 'Frequency', 'fugacity level' = 'fugacity', 'intake fraction' = 'intake',
  'diffusitivity' = 'diffusivity','diffution coefficient' = 'diffusion coefficient',
'body  weight' = 'body weight',  'activity' = 'activity paterns(time)', 'absorption' = 'absorption efficiency',
'trasfer efficiency' = 'transfer efficiency','use patterns' = 'activity paterns(time)', 'layer thickness' = 'thickness',
 'volume fraction' = 'volume', 'building volume' = 'volume','volume/mass' = 'volume',
'consumer products use' = 'Frequency','number of uses' = 'Frequency', 'concentration/area' = 'Concentration',      
   'mass transfer efficiency' = 'transfer efficiency', 'initial age of children' = 'age','active residue' = 'residue level',
   'load pesticide by hand' = 'Amount','mass' = 'Amount', 'ventilation flow' = 'ventilation rate',
   'fraction of the product' = 'fraction of chemical',  'chemical Intake' = 'intake', 'intake time' = 'intake',
 'air volume' = 'volume', 'FoodEx2 EFSA database' = 'Concentration',
                'maximum residue limit (MRL) values (Codex Alimentarius)' = 'residue level',
             'limit of detection' = 'residue level', 'remaining fraction' = 'residue level',
     'pesticide use reports (PUR)' = 'Frequency', 'food ingestion ration' = 'Consumption',
 'reaction  half-lives' = 'half-life values', 'biotransformation half-life' = 'half-life values',
      'half-life in air' = 'half-life values', 'skin laoding' = 'deposition fraction',
     'fraction' = 'fraction of chemical', 'exposure fraction' = 'Duration',
      'indoor exposure fraction' = 'Duration', 'fraction of exposed skin' = 'surface area',
      'air exchange/h' = 'Rate', 'diffusivity' = 'diffusion coefficient',
 'NOAELs for substance s' = 'NOAEL values', 'vaporization percentage' = 'volatilazation fraction'))

##We remove leading and trailing white space on each row
df_Tot_Dm3 <-df_Tot_Dm3 %>% 
  mutate_all(str_trim)


####Joint the Filtered Known models and Data
## that is the dfs df_Tot_Dm3 and f_Models_filt
df_Km_D <-full_join(df_Tot_Dm3, df_Models_filt, by = "Article.ID", multiple = "all", relationship = "many-to-many")
View(df_Km_D)

##### Replace String with Another String on the Known models column
library(dplyr)
df_Km_D <- df_Km_D %>%
  mutate(Knownmod = dplyr::recode(Knownmod, "Monte Carlo Risk Assessment (MCRA) RIVM/EFSA" = "MCRA",
         "Consumer Exposure & Uptake Model (ConsExpo) RIVM" = "ConsExpo",
         "Stoch. Human Exp. Dose Simul. model  (SHEDS-multimedia) USEPA" = "SHEDS",
         'Multimedia transport & transform. model (CalTOX) USEPA' = 'CalTOX',
         'Pesticide Residue Intake Model (PRIMo) EFSA' = 'PRIMo',
         'Targeted Risk Assessment tool (Ecetoc TRA)' = 'Ecetoc',
         'Risk and Exposure Tool (EGRET) ESIG/EU' = 'EGRET',
         'Environment Program (USEtox) model UN' = 'USEtox',
         'Risk Assess. Ident. And Ranking model (RAIDAR) cemc-TrentU CA' = 'RAIDAR',
         'Probabilistic Aggreg. Consm. Exp. Model (PACEM) RIVM' = 'PACEM',
         'Exposure Forecasting (ExpoCast) USEPA' = 'ExpoCast',
         'Modelling Exp. to chemic. for Risk assess. (MERLIN-Expo) EU' = 'MERLIN-Expo'))


####Joint the Filtered Known models and Data
## that is the dfs df_Tot_Dm3 and f_Models_filt
df_Km_D_sh <-full_join(df_Tot_Dm3, df_Models_filt_short, by = "Article.ID", multiple = "all", relationship = "many-to-many") %>% 
  distinct(Article.ID,Data,Knownmod, .keep_all = TRUE) %>% 
  na.omit()

```

Data versus filtered Known models

```{r}
Models_Km_filt_facet <- df_Km_D_sh %>% 
  group_by(Data, Knownmod) %>% 
  summarize(n = n()) %>%
  mutate(perc = 100*n/sum(n)) %>%
  ggplot(aes(x = Data, y = perc, fill = Knownmod)) +
  geom_bar(stat = "identity") +
  # scale_fill_discrete(name="Chemicals") +
  geom_col(show.legend = F) +
  facet_grid( ~ Knownmod, labeller = label_wrap_gen(width=10)) + 
  coord_flip() +
  ylim(c(0, 142)) +
  geom_text(aes(label = round(perc, 0)), size =5.5, hjust = -.1) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(strip.text.x = element_text(face = 'bold', size = 14),
        strip.background = element_rect(fill="white", colour="black",size=1)) + #change the facets font)
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 14),
        axis.text.y = element_text(face = 'bold', size = 16),
        #legend.text = element_text(size = 11),
        #legend.title = element_text(size = 11),
        #legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 20),
        plot.caption = element_text(face = 'bold', size = 10),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Parameters versus Known models", 
       #caption = "ONTOX - Scoping review study",
       x = "Parameters", y = "percentage (%)")
Models_Km_filt_facet
```

# Plot Input variables versus Estimation methodolgies

```{r}
#Join the df df_Tot_Dm3 that contain clean the Data column and dataIV_m that contains Chemicals, #Exposure_route and Est_m data
df_D <- full_join(df_Tot_Dm3, dataIV_m, by = "Article.ID", multiple = "all")
View(df_D)
```

Plot with facets of Input variables vs estimated methodolgies

```{r}
Data_Methods_f <- df_D  %>% 
 distinct(Article.ID,Data,Est_M, .keep_all = TRUE) %>% 
  na.omit() %>%  
  group_by(Data, Est_M) %>% 
  summarize(n = n()) %>%
  mutate(perc = 100*n/sum(n)) %>%
  mutate(Est_M = as.factor(Est_M)) %>% 
  ggplot() +
  scale_fill_brewer(palette = "Set2") +
  geom_bar(aes(x = reorder_within(Data, perc, Est_M),  
               y = perc, fill = Est_M), stat = "identity", show.legend = FALSE) +
  geom_text(aes(x = reorder_within(Data, perc, Est_M), y = perc,
                label = round(perc, 0)), size = 6, hjust = -.1,
            position = position_dodge(width = 1),
            inherit.aes = TRUE) +
  facet_wrap(~ Est_M, scales = "free_y", labeller = label_wrap_gen(width=10)) +
  coord_flip() +
  scale_x_reordered() +
  scale_y_continuous(limits = c(0, 119)) +
  theme_bw() +
 theme(legend.position = "none") +
  theme(strip.text.x = element_text(face = 'bold', size = 16),
        strip.background = element_rect(fill="white", colour="black",size=1)) + #change the facets font)) + #change the facets font)
  theme(axis.title = element_text(size = 22),
        axis.text.x = element_text(face = 'bold', size = 14),
        axis.text.y = element_text(face = 'bold', size = 16),
        #legend.text = element_text(size = 11),
        #legend.title = element_text(size = 11),
        #legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(#title = "Parameters versus Estimated Methodologies", 
       #caption = "ONTOX - Scoping review study",
       x = "Parameters", y = "percentage (%)")
Data_Methods_f
```

# Plot input variablies and Exposure route

```{r}
######Plotting with faceting
#####
Data_ExRout_f <- df_D %>% 
  distinct(Article.ID,Data,ExI, .keep_all = TRUE) %>% 
  na.omit() %>% 
  group_by(Data, ExI) %>% 
  summarize(n = n()) %>%
  mutate(perc = 100*n/sum(n)) %>%
  ggplot(aes(x = Data, y = perc, fill = ExI)) +
  geom_bar(stat = "identity") +
  # scale_fill_discrete(name="Chemicals") +
  geom_col(show.legend = F) +
  facet_grid( ~ ExI,labeller = label_wrap_gen(width=10)) + 
  coord_flip() +
  ylim(c(0, 130)) +
  geom_text(aes(label = round(perc, 1)), hjust = -.1) +
  theme_bw() +
  # theme(legend.position = "none") +
  theme(axis.title = element_text(size = 15),
        axis.text.x = element_text(face = 'bold', size = 10),
        axis.text.y = element_text(face = 'bold', size = 10),
        #legend.text = element_text(size = 11),
        #legend.title = element_text(size = 11),
        #legend.key.size = unit(1, 'cm'),
        plot.title = element_text(size = 15),
        plot.subtitle = element_text(size = 15))+
  labs(title = "Data versus Exposure route", 
       caption = "ONTOX - Scoping review study",
       x = "Input variables", y = "percent (%)")
Data_ExRout_f

```
